\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning on Unsecured Trajectories

### 2023-09-12 paper.tex
Updates from Overleaf

\begin{figure}
     \centering
     \subfloat[Continuous world trajectory \label{fig:ADMM-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 2cm 2cm 2cm]{Example_application}}
     \subfloat[Grid-world trajectory. \label{fig:Grid-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 0.5cm 0cm 2cm]{Kacper_result}}
     \caption{Trajectory design of an example map exploration task for a three-robot system with start locations and destinations fixed. Task are planned in a $8 \times 8$ grid world. Zone 1 is obstacle, Zone 2 and Zone 3 are safe zones. Fig~\ref{fig:Grid-example-application} shows the planned trajectory with co-observation schedule in grid-world. Fig~\ref{fig:ADMM-example-application} shows the trajectory design in continuous world optimized with respect to a map exploration task}
     \label{fig:example-application}
 \end{figure}

@@

The result of the simulation is shown in Fig.\ref{fig:ReachabilitySimulation}. The reachability regions are shown as black ellipsoids, and we can show that all of them have empty intersections with Zone 2 and 3. Notice that no explicit constraints have been enabled between reachability regions and obstacles, because we assume that robots have the basic obstacle avoidance capability and can not go through any hard obstacles. Therefore, the intersections between obstacles and ellipsoids, i.e. the case here between agent 3 and zone 1, are tolerable. All constraints have been satisfied and, to the best of their capability, agents have spread out across the map to perform the best possible exploration task.

@@

To address the feasibility challenge and performance trade-off of the ADMM problem discussed in Section \ref{sec:reachability-discussion}, we propose to form \emph{sub-teams} on each route, and setup additional co-observations both within the sub-team and across different sub-teams. Notice that co-observations across sub-teams (named \emph{cross-trajectory co-observation}) are always preferred when feasible, since they provide better security in potential situations where the entire sub-team is compromised.

In this section, we consider only the high-level path planning, while the assignment of the duties in the team will be performed dynamically, making it more difficult for attackers to successfully plan and execute attacks. Cross-trajectory co-observations allows sub-teams to preserve the optimal unsecured trajectories (as shown in Figure \ref{fig:cross-traj-comparison-set}), as it does not require the entire \emph{sub-team} to maintain a close distance with other teams when co-observation occurs.

@@

Assume we have a total of $N$ robots in the system. We define a partition $\cI=\union_p \cI_p$ of the robots such that robots in \emph{sub-team} $\cI_p$ have the same nominal trajectory. We use the planner introduced in Chapter \ref{chapter:Multi-agent-ADMM} to generate multi-robot trajectories with $N_{p}<N$ routes $\{q_p\}_{p=1}^{N_p}$ without reachability and co-observation constraints, where the state $q_p(t), p \in\{0,\ldots,N_{p}\}$ represents the reference position of the $i$-th sub-team at time $t\in\{0, \dots, t_{e}\}$. Within each sub-team, at least one robot is required to follow the reference trajectory to fulfill the required tasks. Meanwhile, the additional robots can switch from one reference trajectory to another in order to perform co-observation with different sub-teams. We refer to the new type of co-observations as \emph{cross-trajectory co-observation}s. The goal is to find the cross trajectory co-observation plan to secure an unsecured MRS trajectory as with a minimal number of additional robots.

To solve this problem, we model the planned trajectory $\{q_p\}_{p=1}^{N_p}$ as a directed \emph{checkpoint graph} $G_{q}=(V_{q}, E_{q})$. Vertices $V_{q}$ are a key locations in $\{q_p\}$ that are used to perform co-observations. Part of these vertices are \emph{checkpoints} which are selected to guarantee that, if robots deviate and reach forbidden regions, they will miss the scheduled co-observations. Other vertices are selected to connect checkpoints on different trajectories. The set $E_{q} = E_{t} \union E_{c}$ consists of directed edges of two types, trajectory edges $E_{t}$ and cross-trajectory edges $E_{c}$, representing paths which robots can take. The trajectory edges $E_{t}$ represent the reference trajectories, where each edge connects two waypoints on the trajectory of the same sub-team. The cross-trajectory edges $E_{c}$ connect vertices on different reference trajectories, with at least one of the vertices being a checkpoint. These edges represent the potential paths that robots can take to travel and perform co-observations between trajectories of different sub-teams. Similar to the work done by \cite{yu2013multi}, we show that additional robots in the sub-team can be formulated as flows in the checkpoint graph, transforming the co-observation planning problem into a network multi-flow problem that can be solved using general linear programming techniques as specialized solver.

In this chapter, we present a solution to the cross-trajectory co-observation planning problem. This approach consists of two components: constructing the checkpoint graph based on unsecured multi-robot trajectories, which includes the location of security checkpoints and cross-trajectory paths, and the formulation of the network flow problem to solve for the co-observation plan.

\subsection{Rapidly-exploring Random Trees}
In this section, we utilize the \rrtstar{} \cite{karaman2010incremental} algorithm to address the task of determining path feasibility from a single starting location to multiple destination points (reference trajectories of other robots in this application). Our objective here is not to optimize specific constraints but to ascertain the existence of feasible paths. While the ADMM based solver \ref{chapter:ADMM review} presented earlier offers a broader range of constraint handling, the RRT* algorithm is better suited to the problem in this section. RRT*'s efficient exploration of the solution space, coupled with its ability to incorporate obstacle constraints, makes it a fitting choice for our specific task.

As a optimal path planning algorithm, \rrtstar{} returns the shortest paths between an initial location and multiple potential goals. We assume that the generated paths can be travelled in both directions (this is used later in our analysis). 
Key functions from \rrtstar{} that are called by the algorithm in this section are
\begin{description}
\item[$\texttt{Cost}(v)$] This function assigns a non-negative cost (total travel distance in our application) to the unique path from the initial position to $v$ . 
\item[$\texttt{Parent}(v)$] This is a function that maps the vertex $v$ to $v'\in V$ such that $(v',v)\in E$
\end{description}

@@

In this section, we describe in detail how we define and search for security checkpoints and use the checkpoints to find cross-trajectory edges and construct a directed security graph $G_{q}=(V_{q},E_{q})$.

\subsubsection{Checkpoints}\label{sec:security-checkpoint}
Let $q\in\real{nmT}$ be the MRS trajectory found in Chapter \ref{chapter:Multi-agent-ADMM} for $m$ sub-teams with time horizon $T=t_{e}$, with $q_{p}(t)$ be the reference trajectory waypoint for sub-team $p$ at time $t$. A checkpoint $v_i=(q_{i},t_{i})$ is a pair composed of a location $q_{i}=q_{p}(t_{i})$ and a time $t_{i}$ representing a co-observation between a sub-team $\cI_p$'s member and robots either from the same sub-team or different ones. For convenience, let $\cI_{v_{i}}$ represent the corresponding sub-team $v_{i}$ belongs to.

Similar to the requirements of co-observations introduced in Chapter \ref{chapter:Reachability_Constraints}, checkpoints $V_{p}=\{ v_{p_{1}}, \dots ,v_{p_{e}}\} \subset V_{q}$ for a reference trajectory $\{q_{p}(t)\}$ of sub-team $p$ need to guarantee that the reachability region between consecutive checkpoints $\mathcal{E}(q_{p_{i}}, q_{p_{i+1}}, t_{p_{i}},t_{p_{i+1}})$ does not intersect with any of the forbidden areas. For convenience, we denote the union of all forbidden areas as $F$; then the requirement can be written as $\mathcal{E}(q_{p_{i}}, q_{p_{i+1}}, t_{p_{i}},t_{p_{i+1}}) \intersect F = \emptyset$, for every $i$.

We provide a heuristic approach to locate the checkpoints on given trajectories; an optimal solution would likely be NP-hard, while the approach below works well enough for our purpose. We begin by adding the start and end locations $q_{p}(t_{0})$, $q_{p}(t_{e})$ to $V_{q}$, where $t_{0}=0$, and $t_{e}$ is the time horizon. Then, we search for the waypoint $q_{p}(t_{1})$ with the largest $t_{1} \in \{t_{0}, \dots ,t_{2}\}$ such that the reachability ellipsoid between $q_{p}(t_{0})$ and $q_{p}(t_{1})$ has no overlap with any of the forbidden areas, i.e. $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{1}), t_{0},t_{1})\intersect F = \emptyset$. Simultaneously, we search backward to find $q_{p}(t_{3})$ with the smallest $t_{3}\in \{t_{1}, \dots ,t_{2}\} $ that meets the reachability requirement $\mathcal{E}(q_{p}(t_{2}),q_{p}(t_{3}),t_{2},t_{3}) \intersect F = \emptyset$. $q_{p}(t_{1})$ and $q_{p}(t_{3})$ are then added to $V_{q}$. Afterwards, we set $t_{0}=t_{1}$, $t_{2}=t_{3}$ and repeat the same process. The search is stopped when $t_{0}>t_{2}$, or when $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{2}),t_{0},t_{2}) \intersect F = \emptyset$. A toy example is shown in Figure \ref{fig:checkpoint-generate}.

@@

\subsubsection{Cross-trajectory edges}\label{sec:cross-traj-edges}
After locating checkpoints $\union_p V_p$ for all sub-teams, we can search for cross-trajectory edges to connected all checkpoints to trajectories. Cross-trajectory edges represents feasible paths between two references trajectories that allow robots to reach a different reference trajectories and perform co-observations with the corresponding sub-teams. Thus, at least one end of the edges must be a security checkpoint $v_{p} \in \union_p V_p$. Additionally, assuming that only one robot is sent at a time between trajectories, the cross-trajectory edges must satisfy the reachability constraints from Chapter~\ref{chapter:Reachability_Constraints} to ensure that no deviation to forbidden areas can be performed while switching trajectories.

@@

To find cross-trajectory edges, we base our approach on RRT*. First, we search for feasible paths between each security checkpoint in each reference trajectory and all other trajectories, as shown in Figure \ref{fig:RRT*}. Since paths returned by RRT* are optimal and bidirectional, we can analyze the travel time and reachability for each candidate cross-trajectories, and  add these edges to the checkpoint graph if they are feasible.

More vigorously, After a fixed iterations of RRT*, we find all possible and close to optimal paths between one security checkpoint $v_{p}=(q_{p},t_{p}) \in V_{p}$ for sub-team $\cI_p$, and $V_{r} = \{ (q_{{r}_{i}}, t_{{r}_{i}}) \}$ where  $q_{{r}_{i}} = q_{r}(t_{{r}_{i}})$ are the entire waypoints on a reference trajectories for a different sub-team $\cI_{r}$ for a fixed number of RRT* iterations. We can calculate the minimal travel time $t=\texttt{Cost}(q)/v_{max}$ for a robot to traverse each path. We call all nodes $v_{{r}_{i}} \in V_{r}$ \emph{arrival node}s if there exist a path between $v_{{r}_{i}}$ and $v_{p}$ with $t_{p}+t<t_{{r}_{i}}$, and the reachability ellipsoid $\mathcal{E}(q_{r}(t_{{r}_{i}}),q_{p},t_{{r}_{i}},t_{p}) \intersect F = \emptyset$. Similarly, \emph{departure node}s are defined for $v_{{r}_{i}}$ with $t_{p}>t_{{r}_{i}}+t$ and $\mathcal{E}(q_{p}, q_{r}(t_{{r}_{i}}),t_{p}, t_{{r}_{i}}) \intersect F = \emptyset$. Given a point $v_{p} \in V_{p}$, we define the \emph{latest departure node} $q_{r}(t^{d}_r)$, and the \emph{earliest arrival node} $q_{r}(t^{a}_r)$ as follows:

>>>>

\begin{figure}
     \centering
     \subfloat[Continuous world trajectory \label{fig:ADMM-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 2cm 2cm 2cm]{Example_application}}
     \subfloat[Grid-world trajectory. \label{fig:Grid-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 0.5cm 0cm 2cm]{Kacper_result}}
     \caption{Trajectory design of an example map exploration task for a three-robot system with start locations and destinations fixed. Task are planned in a $8 \times 8$ grid world. Zone 1 is obstacle, Zone 2 and Zone 3 are safe zones. Fig~\ref{fig:Grid-example-application} shows the planned trajectory with co-observation schedule in grid-world. Fig~\ref{fig:ADMM-example-application} shows the trajectory design in continuous world optimized with respect to a map exploration task}
     \label{fig:example-application}
 \end{figure}
 \rtron{Make sure we refer back to Example 1 when makes sense}

@@

The result of the simulation is shown in Fig.~\ref{fig:ReachabilitySimulation}. The reachability regions are shown as black ellipsoids, and we can show that all of them have empty intersections with Zone 2 and 3. Notice that no explicit constraints have been enabled between reachability regions and obstacles, because we assume that robots have the basic obstacle avoidance capability and can not go through any hard obstacles. Therefore, the intersections between obstacles and ellipsoids, i.e. the case here between agent 3 and zone 1, are tolerable. All constraints have been satisfied and, to the best of their capability, agents have spread out across the map to perform the best possible exploration task.

@@

To address the feasibility challenge and performance trade-off of the ADMM problem discussed in \Cref{sec:reachability-discussion}, we propose to form \emph{sub-teams} on each route, and setup additional co-observations both within the sub-team and across different sub-teams \rtronref{refer to introductory figure}. 
These \emph{cross-trajectory co-observations} allow sub-teams to preserve the optimal unsecured trajectories (as shown in Figure \ref{fig:cross-traj-comparison-set}), as it does not require the entire \emph{sub-team} to maintain a close distance with other teams when co-observation occurs. 
Note that \emph{cross-trajectory co-observation}, are always preferred when feasible, since they provide better security in potential situations where the entire sub-team is compromised. \rtronref{(More discussion in Sec. ??)}.

\rtronhl{Move to discussion}{In this section, we consider only the high-level path planning, while the assignment of the duties in the team will be performed dynamically, making it more difficult for attackers to successfully plan and execute attacks.}

@@

\rtron{Introduce $n_p$ in the previous Section.}

In \Cref{sec:??} we assumed that the number of robot is equal to the number of routes $n_p$.
In this section, we  assume that we have a total of $n>n_p$ robots  available. \rtronhl{Clarify: since the robots change teams, this partition is time-varying, or can we get by without defining this partition explicitly?} We define a partition $\cI=\union_p \cI_p$ of the robots such that robots in \emph{sub-team} $\cI_p$ have the same nominal trajectory. We use the planner introduced in  \Cref{chapter:Multi-agent-ADMM} to generate multi-robot trajectories with $n_p$ routes $\{q_p\}_{p=1}^{N_p}$ without reachability and co-observation constraints, where the state $q_p(t), p \in\{0,\ldots,N_{p}\}$ represents the reference position of the $i$-th sub-team at time $t\in\{0, \dots, t_{e}\}$. \rtron{Again, make sure notation is consistent}
Within each sub-team, at least one robot is required to follow the reference trajectory to fulfill the required tasks. Meanwhile, the additional robots can switch from one reference trajectory to another in order to perform co-observation with different sub-teams. 
We refer to the new type of co-observations as \emph{cross-trajectory co-observations}. The goal is to find the cross trajectory co-observation plan to secure an unsecured MRS trajectory with a minimal number of additional robots.

\rtronhl{This can be summarized more, since it is discussed later in this same page}{To solve this problem, we model the planned trajectory $\{q_p\}_{p=1}^{N_p}$ as a directed \emph{checkpoint graph} $G_{q}=(V_{q}, E_{q})$. 
Vertices $V_{q}$ are a key locations in $\{q_p\}$ that are used to perform co-observations. Part of these vertices are \emph{checkpoints} which are selected to guarantee that, if robots deviate and reach forbidden regions, they will miss the scheduled co-observations. 
Other vertices are selected to connect checkpoints on different trajectories.
The set $E_{q} = E_{t} \union E_{c}$ consists of directed edges of two types, trajectory edges $E_{t}$ and cross-trajectory edges $E_{c}$, representing paths which robots can take. The trajectory edges $E_{t}$ represent the reference trajectories, where each edge connects two waypoints on the trajectory of the same sub-team. The cross-trajectory edges $E_{c}$ connect vertices on different reference trajectories, with at least one of the vertices being a checkpoint. These edges represent the potential paths that robots can take to travel and perform co-observations between trajectories of different sub-teams.} Similar to the work done by \cite{yu2013multi}, we show that additional robots in the sub-team can be formulated as flows in the checkpoint graph, transforming the co-observation planning problem into a network multi-flow problem that can be solved using general linear programming techniques as specialized solver.

This section presents the two components of our approach: constructing the checkpoint graph based on unsecured multi-robot trajectories, and the formulation and solution of the network flow problem.

\subsection{Rapidly-exploring Random Trees}
In this section, we use the \rrtstar{} \cite{karaman2010incremental} algorithm to find paths from a single starting location to multiple destination points (i.e., the reference trajectories of other robots, in our method). \rtronhl{Move to the end of this RRT* section}{Our objective is not to optimize specific constraints but to ascertain the existence of feasible paths. While the ADMM based solver \Cref{chapter:ADMM review} presented earlier offers a broader range of constraint handling, the RRT* algorithm is better suited to the problem in this section. RRT*'s efficient exploration of the solution space, coupled with its ability to incorporate obstacle constraints, makes it a fitting choice for building the checkpoint graph.}

As a optimal path planning algorithm, \rrtstar{} returns the shortest paths between an initial location and points in the free configuration space, organized as a tree. We assume that the generated paths can be travelled in both directions (this is used later in our analysis). 
Key functions from \rrtstar{} that are also used during constructions of the checkpoint graph are
\begin{description}
\item[$\texttt{Cost}(v)$] This function assigns a non-negative cost (total travel distance in our application) to the unique path from the initial position to $v$ . 
\item[$\texttt{Parent}(v)$] This is a function that maps the vertex $v$ to $v'\in V$ such that $(v',v)\in E$.
\end{description}

@@

In this section, we describe in detail how we define and search for security checkpoints and how to use \rrtstar{} to construct the checkpoint graph $G_{q}$.

\subsubsection{Checkpoints}\label{sec:security-checkpoint}
Let $q\in\real{nmT}$ be the MRS trajectory found in Chapter \ref{chapter:Multi-agent-ADMM} for $m$ sub-teams with time horizon $T=t_{e}$, with $q_{p}(t)$ be the reference trajectory waypoint for sub-team $p$ at time $t$. A checkpoint $v_i=(q_{i},t_{i})$ is a pair composed of a location $q_{i}=q_{p}(t_{i})$ and a time $t_{i}$ representing a co-observation between a sub-team $\cI_p$'s member and robots either from the same sub-team or different ones. For convenience, let $\cI_{v_{i}}$ represent the sub-team $v_{i}$ belongs to.

Similar to the requirements of co-observations introduced in  \Cref{chapter:Reachability_Constraints}, \rtronhl{Consider wrapping this in definition environment} {we define checkpoints $V_{p}=\{ v_{p_{1}}, \dots ,v_{p_{e}}\} \subset V_{q}$ for a reference trajectory $\{q_{p}(t)\}$ of sub-team $p$ in a way that guarantees that the reachability region between consecutive checkpoints $\mathcal{E}(q_{p_{i}}, q_{p_{i+1}}, t_{p_{i}},t_{p_{i+1}})$ does not intersect with any of the forbidden areas.}  The requirement can be formally stated as $\mathcal{E}(q_{p_{i}}, q_{p_{i+1}}, t_{p_{i}},t_{p_{i+1}}) \intersect F = \emptyset$ for every $i$, where $F$ is the union of all forbidden areas as $F$.

We provide a heuristic approach to locate the checkpoints on given trajectories (an optimal solution would likely be NP-hard, while the approach below works well enough for our purpose). 
\rtron{State this as an Algorithm?} We begin by adding the start and end locations $q_{p}(t_{0})$, $q_{p}(t_{e})$ to $V_{q}$, where $t_{0}=0$, and $t_{e}$ is the time horizon. Then, we search for the waypoint $q_{p}(t_{1})$ with the largest $t_{1} \in \{t_{0}, \dots ,t_{2}\}$ such that the reachability ellipsoid between $q_{p}(t_{0})$ and $q_{p}(t_{1})$ has no overlap with any of the forbidden areas, i.e. $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{1}), t_{0},t_{1})\intersect F = \emptyset$. Simultaneously, we search backward to find $q_{p}(t_{3})$ with the smallest $t_{3}\in \{t_{1}, \dots ,t_{2}\} $ that meets the reachability requirement $\mathcal{E}(q_{p}(t_{2}),q_{p}(t_{3}),t_{2},t_{3}) \intersect F = \emptyset$. $q_{p}(t_{1})$ and $q_{p}(t_{3})$ are then added to $V_{q}$. Afterwards, we set $t_{0}=t_{1}$, $t_{2}=t_{3}$ and repeat the same process. The search is stopped when $t_{0}>t_{2}$, or when $\mathcal{E}(q_{p}(t_{0}),q_{p}(t_{2}),t_{0},t_{2}) \intersect F = \emptyset$. A toy example is shown in Figure \ref{fig:checkpoint-generate}.

@@

\subsubsection{Cross-trajectory edges}\label{sec:cross-traj-edges}
After locating checkpoints $\union_p V_p$ for all sub-teams, we can search for cross-trajectory edges to connected all checkpoints to trajectories. Cross-trajectory edges represents feasible paths between two references trajectories that allow robots to reach a different reference trajectories and perform co-observations with the corresponding sub-teams. 
Thus, at least one end of the edges must be a security checkpoint $v_{p} \in \union_p V_p$. Additionally, assuming that only one robot is sent at a time between trajectories, the cross-trajectory edges must satisfy the reachability constraints from~\Cref{chapter:Reachability_Constraints} to ensure that no deviation to forbidden areas can be performed while switching trajectories.

@@

To find cross-trajectory edges, we first find trees of feasible paths using \rrtstar{} and position information alone; we then prune these trees by considering time constraints (that consider the time needed to physically travel from one trajectory to the other while meeting other robots at the two endpoints) and ellipsoidal rechability constraints.

We search for feasible paths between each security checkpoint and all other trajectories (ignoring, for the moment, any timing constraint), as shown in \Cref{fig:RRT*} \rtron{Missing figure}. 
%Paths returned by RRT* are optimal and bidirectional.%, we can analyze the travel time and reachability for each candidate cross-trajectories, and  add these edges to the checkpoint graph if they are feasible.
More precisely, after a fixed number of  iterations, \rrtstar{}  finds all feasible, quasi-optimal paths between one security checkpoint $v_{p}=(q_{p},t_{p}) \in V_{p}$ for sub-team $\cI_p$, and $V_{r} = \{ (q_{{r}_{i}}, t_{{r}_{i}}) \}$ where  $q_{{r}_{i}} = q_{r}(t_{{r}_{i}})$ are the entire waypoints on a reference trajectories for a different sub-team $\cI_{r}$.
We can calculate the minimal travel time $t=\texttt{Cost}(q)/v_{max}$ for a robot to traverse each path. We call all nodes $v_{{r}_{i}} \in V_{r}$ \emph{arrival node}s if there exist a path between $v_{{r}_{i}}$ and $v_{p}$ with $t_{p}+t<t_{{r}_{i}}$, and the reachability ellipsoid $\mathcal{E}(q_{r}(t_{{r}_{i}}),q_{p},t_{{r}_{i}},t_{p}) \intersect F = \emptyset$. Similarly, \emph{departure node}s are defined for $v_{{r}_{i}}$ with $t_{p}>t_{{r}_{i}}+t$ and $\mathcal{E}(q_{p}, q_{r}(t_{{r}_{i}}),t_{p}, t_{{r}_{i}}) \intersect F = \emptyset$. Given a point $v_{p} \in V_{p}$, we define the \emph{latest departure node} $q_{r}(t^{d}_r)$, and the \emph{earliest arrival node} $q_{r}(t^{a}_r)$ as follows:
